Succinct representations for abstract interpretation

### 2011-11-04 papers/PLDI2012/implicitization.tex
typos

\begin{abstract}
Abstract interpretation techniques can be made more precise by distinguishing paths inside loops, at the expensive of possibly exponential complexity. SMT-solving techniques and sparse representations of paths and sets of paths avoid this pitfall.

@@

There are many possible combinations of the above techniques, and it is not evident which ones perform more or less precisely or more or less efficiency on real-life examples. One needs to experiment.
Unfortunately, the published literature on the subject lacks experimental comparative assessments. One purpose of this article is therefore to propose such experimental results.

@@

Once an inductive invariant $\gamma((\abstr{x}_p)_{p \in P})$ has been obtained, one can attempt \emph{decreasing} or \emph{narrowing} iterations to reduce it. In their simplest form, this just means running the following operation until a fixpoint or a maximal number of iterations are reached: for any $p'$, replace $\abstr{x}_{p'}$ by $\abstr{x}_p \cap \left(\bigsqcup_{p \in P} \abstr{\tau}_{p,p'} (\abstr{x}_p)\right)$. The result also defines an inductive invariant. These decreasing iterations are indispensable to recover properties from guards (tests) in the program in most iteration settings; unfortunately, certain loops, particularly those involving identity (no-operation) transitions, may foil them: the iterations immediately reach a fixpoint and do not decrease further (see example in \S\ref{subsec:rate_lim}). Sections \ref{sec:guided} and \ref{sec:path_focusing} describe techniques that work around this problem.

@@

\begin{figure}
\label{fig:multigraph}
\centering
\begin{minipage}[c]{.19\textwidth}
\begin{tikzpicture}[->,>=stealth',auto,node distance=1.1cm,
                    semithick,font=\footnotesize]

@@

  \path [transition] 
		(n0) edge  [out=0, in=0]            node {} (n9);
  \path [transition] 
		(n0) edge  [out=180, in=180]        node {} (n9);
  \path [transition] 
		(n0) edge  [out=205, in=155]            node {} (n9);
  \path [transition] 
		(n0) edge  [out=230, in=130]            node {} (n9);
  \path [transition] 
		(n0) edge  [out=255, in=105]            node {} (n9);
  \path [transition] 
		(n0) edge  [out=-25, in=25]            node {} (n9);
  \path [transition] 
		(n0) edge  [out=-50, in=50]            node {} (n9);
  \path [transition] 
		(n0) edge  [out=-75, in=75]            node {} (n9);
\end{tikzpicture}
\end{minipage}
\caption{Expansion of the transition graph into a multigraph.}
\end{figure}

>>>>

\begin{abstract}
Abstract interpretation techniques can be made more precise by distinguishing
paths inside loops, at the expense of possibly exponential complexity. SMT-solving techniques and sparse representations of paths and sets of paths avoid this pitfall.

@@

There are many possible combinations of the above techniques, and it is not
evident which ones perform more or less precisely or more or less efficiently on real-life examples. One needs to experiment.
Unfortunately, the published literature on the subject lacks experimental comparative assessments. One purpose of this article is therefore to propose such experimental results.

@@

Once an inductive invariant $\gamma((\abstr{x}_p)_{p \in P})$ has been obtained,
one can attempt \emph{decreasing} or \emph{narrowing} iterations to reduce it.
In their simplest form, this just means running the following operation until a
fixpoint or a maximal number of iterations are reached: for any $p'$, replace
$\abstr{x}_{p'}$ by $\abstr{x}_{p'} \cap \left(\bigsqcup_{p \in P}
\abstr{\tau}_{p,p'} (\abstr{x}_p)\right)$. The result also defines an inductive
invariant. These decreasing iterations are indispensable to recover properties
from guards (tests) in the program in most iteration settings; unfortunately,
certain loops, particularly those involving identity (no-operation) transitions,
may foil them: the iterations immediately reach a fixpoint and do not decrease
further (see example in \S\ref{subsec:rate_lim}). Sections \ref{sec:guided} and
\ref{sec:path_focusing} describe techniques that work around this problem.

@@

\begin{figure}
\centering
\begin{minipage}[c]{.19\textwidth}
\begin{tikzpicture}[->,>=stealth',auto,node distance=1.1cm,
                    semithick,font=\footnotesize]

@@

  \path [transition] 
		(n0) edge  [out=0, in=0]            node {} (n9);
  \path [transition] 
		(n0) edge  [out=180, in=180]        node {} (n9);
  \path [transition] 
		(n0) edge  [out=205, in=155]            node {} (n9);
  \path [transition] 
		(n0) edge  [out=230, in=130]            node {} (n9);
  \path [transition] 
		(n0) edge  [out=255, in=105]            node {} (n9);
  \path [transition] 
		(n0) edge  [out=-25, in=25]            node {} (n9);
  \path [transition] 
		(n0) edge  [out=-50, in=50]            node {} (n9);
  \path [transition] 
		(n0) edge  [out=-75, in=75]            node {} (n9);
\end{tikzpicture}
\end{minipage}
\caption{Expansion of the transition graph into a multigraph.}
\label{fig:multigraph}
\end{figure}
